---
title: "BS_figures"
author: "Dr. Burchfield"
output: html_document
---

# Load results and data

```{r message=F, warning=F}
library(ggpubr)
library(RColorBrewer)
library(knitr)

source("BS_load.R")
source("BS_func.R")

mcorn <- readRDS("./out/models/mf_corn.RDS")
dcorn <- readRDS("./out/models/df_corn.RDS")

msoy <- readRDS("./out/models/mf_soy.RDS")
dsoy <- readRDS("./out/models/df_soy.RDS")

mwwheat <- readRDS("./out/models/mf_wwheat.RDS")
dwwheat <- readRDS("./out/models/df_wwheat.RDS")

malfalfa <- readRDS("./out/models/mf_alfalfa.RDS")
dalfalfa <- readRDS("./out/models/df_alfalfa.RDS")

mhay <- readRDS("./out/models/mf_hay.RDS")
dhay <- readRDS("./out/models/df_hay.RDS")

#https://www.nature.com/articles/nmeth.1618
wong_colors <- c("#E69F00", "#e34234", "#009E73", "#F0E442", "#0072B2")
crops <- as.factor(c("corn", "soy", "wwheat", "alfalfa", "hay"))
clean_crops <- as.factor(c("Corn", "Soy", "Winter wheat", "Alfalfa", "Hay"))
names(wong_colors) <- clean_crops
colors_f <- scale_fill_manual(name = "CROPS", values = wong_colors)
colors_c <- scale_color_manual(name = "CROPS", values = wong_colors)

mlist <- list(mcorn, msoy, mwwheat, malfalfa, mhay)

dflist <- list()

for (i in 1:length(crops)) {
  
  crop <- crops[[i]]
  out <- prep_data(crop, scale = T)
  county_sub <- out[[2]] 
  df <- prep_attr_data(crop, scale = F)
  st_geometry(df) <- NULL
  df$CROP <- crop
  df$CCROP <- clean_crops[[i]]
  df$PROP_CROP <- df$PERC_CROP/100
  dflist[[i]] <- df
  
}

df <- do.call("rbind.data.frame", dflist)
df <- merge(county, df, by = "GEOID", all = T)

df$BS <- ifelse(df$DIFF > mean(df$DIFF, na.rm=T) + 1*sd(df$DIFF, na.rm=T), "> 1 sd", NA)
df$BS[df$DIFF < mean(df$DIFF, na.rm=T) - 1*sd(df$DIFF, na.rm=T)] <- "< 1 sd"
df$BS[is.na(df$DIFF)] <- "Average spot"

race <- readRDS("./out/race.RDS") %>% filter(YEAR %in% c(2012, 2017))
race <- race %>% group_by(GEOID) %>% summarize_all(funs(mean), na.rm=T)
race[race == "NaN"] <- NA

df$PROP_CROP[is.na(df$PROP_CROP)] <- 0
df2 <- merge(df, race, by = "GEOID", all = T)
df2$WD <-df2$DIFF * df2$PROP_CROP

# ACROSS crop bright spots
dfs <- df2 %>%
  group_by(GEOID) %>%
  summarize(N = sum(PROP_CROP, na.rm=T),
    #N = length(DIFF[!is.na(DIFF)]),
            TOTAL = sum(WD, na.rm=T)) %>%
  mutate(WAVG = TOTAL/N)
dfa <-  df2 %>%
    select(-c(geometry, WD, PROP_CROP, CROP, CCROP, Z, DIFF, STATEFP, FRR.x, FRR.y)) %>% # remove null model attr
    group_by(GEOID) %>% 
    summarize_all(funs(mean), na.rm=T)
st_geometry(dfa) <- NULL
dfa[dfa == "NaN"] <- NA

dfs <- merge(dfs, dfa, by = "GEOID", all = T)
dfs$BS <- ifelse(dfs$WAVG > mean(dfs$WAVG, na.rm=T) + 1*sd(dfs$WAVG, na.rm=T), "> 1 sd", NA)
dfs$BS[dfs$WAVG < mean(dfs$WAVG, na.rm=T) - 1*sd(dfs$WAVG, na.rm=T)] <- "< 1 sd"
dfs$BS[is.na(dfs$BS)] <- "Average spot"
dfs$BS <- as.factor(dfs$BS)

library(reshape2)
library(scales)

# scale dfs
dfscale <- dfs %>% select(-c(GEOID, BS, N, TOTAL, WAVG))
st_geometry(dfscale) <- NULL
dfscale <- as.data.frame(scale(dfscale)) 
dfscale <- cbind.data.frame(dfs %>% select(c(GEOID, BS, N, TOTAL, WAVG)), dfscale)
```

***

# Figure 1: Nonlinear yield responses

```{r}
t <- prep_data("corn", scale=F)[[1]]
mgdd <- mean(t$GDD, na.rm=T)
sgdd <- sd(t$GDD, na.rm=T)
mtp <- mean(t$TP, na.rm=T)
stp <- sd(t$TP, na.rm=T)

crops <-crops[!crops %in% c("sorghum", "cotton")]
clean_crops <- clean_crops[!clean_crops %in% c("Sorghum", "Cotton")]

plot_nle <- function(variable, m, s) {
  
  out_list <- list()
  for (i in 1:length(crops)) {

    df <- mlist[[i]]$summary.random[[variable]][,c(1,4:6)]
    df$CROPS <- crops[[i]]
    df$CROP_NAME <- clean_crops[[i]]
    out_list[[i]] <- df
    
  }
  nle <- do.call("rbind.data.frame", out_list)
  nle$ID2 <- nle$ID
  nle$ID <- (nle$ID * s) + m
  
  plt <- ggplot(data = nle) + 
    geom_line(aes(x = ID, y = `0.5quant`, group = CROP_NAME)) +
    geom_ribbon(aes(x = ID, y = `0.5quant`, ymin = `0.025quant`, ymax = `0.975quant`, fill = CROP_NAME), alpha = 0.7) +
    project_theme +
    colors_f +
    labs(x = variable, 
         y = "Standardized yield effect") +
    theme(legend.title = element_blank())
  
  return(plt)
  
}

library(ggpubr)
gdd <- plot_nle("GDD", mgdd, sgdd) + xlab("Growing degree days (°C)") + theme(text = element_text(size = 20),
                                                                              axis.text = element_text(size = 18)
                                                                              ) + ylim(c(-2, 1))+ theme(legend.text=element_text(size=18))
tp <- plot_nle("TP", mtp, stp) + xlab("Total precipitation (mm)") + ylab("") + theme(text = element_text(size = 20),
                                                                              axis.text = element_text(size = 18),
                                                                                     legend.position = "none") + ylim(c(-2,1)) +  theme(legend.text=element_text(size=18))

ggarrange(gdd, tp, labels = c("A", "B"), ncol = 2, nrow = 1, common.legend = T, legend = "bottom", font.label = list(size = 24))
ggsave("./figs/nonlinear_climate.png", plot = last_plot(), width = 18, height = 8)
```


# Figure 2: Regression coefficients 

```{r echo=F}
plot_all_coef <- function(mlist) {
  
  out <- list()
  
  for (i in 1:length(mlist)) {
    
    model <- mlist[[i]]
    coef <- model$summary.fixed[,c(1, 3, 5)] # mean, 0.025, 0.985
    colnames(coef) <- c("MEAN", "LOW", "HIGH")
    coef$VARS <- rownames(coef)
    coef <- coef %>% filter(VARS != "(Intercept)")
    # drop year factors for now and intercept
    # coef <- coef[2:9,]
    coef$CROP <- clean_crops[i]
    out[[i]] <- coef
  
    
    
  }
  
  out <- do.call("rbind.data.frame", out)
  out$VOI <- NA
  out$VOI[out$VARS == "PERC_IRR"] <- "% ag. land irrigated"
  out$VOI[out$VARS == "PERC_AG"] <- "% agricultural land"
  out$VOI[out$VARS == "T_CEC_SOIL"] <- "Cation exchange capacity"
  out$VOI[out$VARS == "T_OC"] <- "Topsoil organic carbon"
  out$VOI[out$VARS == "T_PH_H2O"] <- "Topsoil pH"
  out$VOI[out$VARS == "T_ESP"] <- "Topsoil sodicity"
  out$VOI[out$VARS == "factor(YEAR)2009"] <- "y2009"
  out$VOI[out$VARS == "factor(YEAR)2010"] <- "y2010"
  out$VOI[out$VARS == "factor(YEAR)2011"] <- "y2011"
  out$VOI[out$VARS == "factor(YEAR)2012"] <- "y2012"
  out$VOI[out$VARS == "factor(YEAR)2013"] <- "y2013"
  out$VOI[out$VARS == "factor(YEAR)2014"] <- "y2014"
  out$VOI[out$VARS == "factor(YEAR)2015"] <- "y2015"
  out$VOI[out$VARS == "factor(YEAR)2016"] <- "y2016"
  out$VOI[out$VARS == "factor(YEAR)2017"] <- "y2017"
  out$VOI[out$VARS == "factor(YEAR)2018"] <- "y2018"
  
  ggplot(out %>% filter(!VARS %in% c("(Intercept)"), 
                        !VOI %in% c("y2009", "y2010", "y2011", "y2012", "y2013", "y2014", "y2015", "y2016", "y2017", "y2018"))) +
    geom_errorbarh(height = 0, aes(xmin = LOW, xmax = HIGH, 
                                   y = reorder(VOI, LOW), color = CROP), position = position_dodge(width = 0.3)) +
    project_theme +
    geom_point(aes(x = MEAN, y = reorder(VOI, LOW), color = CROP), size = 2, alpha = 0.8, 
               position = position_dodge(width = 0.3)) +
    geom_vline(xintercept = 0) +
    labs(x = "",
         y = "",
         color = "") +
    colors_c +
  theme(legend.position = "right",
        legend.title = element_blank()) 
  
  
}

coef <- plot_all_coef(mlist) + theme(text = element_text(size = 12),
                                     axis.text = element_text(size = 12),
                                     legend.text = element_text(size=12))
# ggsave("./figs/coefficients_notime.png", width = 7, height = 3)
```

```{r echo=F}
plot_all_coef <- function(mlist) {
  
  out <- list()
  
  for (i in 1:length(mlist)) {
    
    model <- mlist[[i]]
    coef <- model$summary.fixed[,c(1, 3, 5)] # mean, 0.025, 0.985
    colnames(coef) <- c("MEAN", "LOW", "HIGH")
    coef$VARS <- rownames(coef)
    coef <- coef %>% filter(VARS != "(Intercept)")
    # drop year factors for now and intercept
    # coef <- coef[2:9,]
    coef$CROP <- clean_crops[i]
    out[[i]] <- coef
  
    
    
  }
  
  out <- do.call("rbind.data.frame", out)
  out$VOI <- NA
  out$VOI[out$VARS == "PERC_IRR"] <- "% ag. land irrigated"
  out$VOI[out$VARS == "PERC_AG"] <- "% agricultural land"
  out$VOI[out$VARS == "T_CEC_SOIL"] <- "Cation exchange capacity"
  out$VOI[out$VARS == "T_OC"] <- "Topsoil organic carbon"
  out$VOI[out$VARS == "T_PH_H2O"] <- "Topsoil pH"
  out$VOI[out$VARS == "T_ESP"] <- "Topsoil salinity"
  out$VOI[out$VARS == "factor(YEAR)2009"] <- "2009"
  out$VOI[out$VARS == "factor(YEAR)2010"] <- "2010"
  out$VOI[out$VARS == "factor(YEAR)2011"] <- "2011"
  out$VOI[out$VARS == "factor(YEAR)2012"] <- "2012"
  out$VOI[out$VARS == "factor(YEAR)2013"] <- "2013"
  out$VOI[out$VARS == "factor(YEAR)2014"] <- "2014"
  out$VOI[out$VARS == "factor(YEAR)2015"] <- "2015"
  out$VOI[out$VARS == "factor(YEAR)2016"] <- "2016"
  out$VOI[out$VARS == "factor(YEAR)2017"] <- "2017"
  out$VOI[out$VARS == "factor(YEAR)2018"] <- "2018"
  
  outs <- out %>% filter(VOI %in% c("2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"))
  outs$VOI <- as.numeric(as.character(outs$VOI))
  
  ggplot(outs, aes(x = VOI, y = MEAN, group = CROP, color = CROP)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin = LOW, ymax = HIGH), width = 0.2) +
    project_theme +
    labs(x = "",
         y = "Mean estimated effect",
         color = "")+
    colors_c +
  theme(legend.position = "right",
        legend.title = element_blank()) +
    scale_x_continuous(breaks = 2009:2018, labels = 2009:2018)

}

tcoef <- plot_all_coef(mlist)  + theme(text = element_text(size = 12),
                                     axis.text = element_text(size = 12),
                                     legend.text = element_text(size=12))

# ggsave("./figs/coefficients_time_only.png", width = 10, height = 5)
```

```{r}
ggarrange(coef, tcoef, labels = c("A", "B"), common.legend = T, nrow = 2, ncol =1, legend = "bottom")
ggsave("./figs/regression_results.png", width = 8, height = 8)
```

***

# Figure 3: Crop-specific differences

```{r echo=F, message=F, warning=F}
dfus <- df  # for crop specific
dfus$DIFF[dfus$DIFF > 4] <- 4
dfus$DIFF[dfus$DIFF < -4] <- -4

limit <- max(abs(dfus$DIFF), na.rm=T) * c(-1, 1)

corn <- ggplot() +
  geom_sf(data = dfus %>% filter(CROP == "corn"), color = "transparent", size = 0.05, aes(fill = DIFF)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
     scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  labs(fill = "",
       title = "C. Corn",
       subtitle = "") + theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

soy <- ggplot() +
  geom_sf(data = dfus %>% filter(CROP == "soy"), color = "transparent", size = 0.05, aes(fill = DIFF)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
     scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  labs(fill = "",
       title = "D. Soy",
       subtitle = "") + theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

wwheat <- ggplot() +
  geom_sf(data = dfus %>% filter(CROP == "wwheat"), color = "transparent", size = 0.05, aes(fill = DIFF)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
    scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  labs(fill = "",
       title = "E. Winter wheat",
       subtitle = "") + theme(text = element_text(size = 20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

alfalfa <- ggplot() +
  geom_sf(data = dfus %>% filter(CROP == "alfalfa"), color = "transparent", size = 0.05, aes(fill = DIFF)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
    scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", 
                         midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  labs(fill = "",
       title = "A. Alfalfa",
       subtitle = "") + theme(text = element_text(size = 20), 
                              legend.text = element_text(size = 20),
                              legend.position = "right") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank()) 

hay <- ggplot() +
  geom_sf(data = dfus %>% filter(CROP == "hay"), color = "transparent", size = 0.05, aes(fill = DIFF)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
     scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  labs(fill = "",
       title = "B. Hay",
       subtitle = "") + 
  theme(text = element_text(size = 20) 
        ) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

get_legend<-function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

limit <- max(abs(dfs$WAVG), na.rm=T) * c(-1, 1)
wd <- ggplot() +
  geom_sf(data = dfs, color = "transparent", size = 0.05, aes(fill = WAVG)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
  scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  labs(fill = "",
       title = "",
       subtitle = "") +
  labs(fill = "",
       title = "F. Weighted average",
       subtitle = "") + theme(text = element_text(size = 20), legend.text = element_text(size = 20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

lout <- get_legend(alfalfa) #+ guides(shape = guide_legend(override.aes = list(size=40)))

ggarrange(alfalfa, hay, corn, soy, wwheat, wd, ncol = 2, nrow = 3, legend.grob = lout, common.legend = T, legend = "right")

ggsave("./figs/diff_plot_crop.png", width = 15 , height= 15)

```

***

# Figure 4: Box and whisker for WA

Sum expenses:

```{r eval=F}
medians <- dfs %>% 
  select(-c(WAVG, N, GEOID, TOTAL)) %>%
  group_by(BS) %>% summarize_all(funs(median), na.rm=T)

inputhi <- medians %>% group_by(BS) %>%
  summarize(TOTAL = fert + chem + labor_expense + machinery)
```

```{r}
# dfs %>% group_by(BS) %>% summarize(med = median(insur_acres, na.rm=T))
plt <- melt(dfscale, id.vars = c("BS"),
            measure.vars = c("fert", "chem", "machinery", "labor_expense", "gvt_prog", "insur_acres",
                             "exp", "occup", "tenant", "acres_per_op", "BLACKP", "HISPP", "WHITEP", "female",  "ED", "SDI", "PNC", "AREA_MN_AG")) %>%
  filter(BS != "Average")
c <- plt
c$VOI <- as.character(plt$variable)
c$CAT <- NA

c$VOIC[c$VOI == "female"] <- "% female operators"
c$CAT[c$VOI == "female"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "BLACKP"] <- "% Black acres operated"
c$CAT[c$VOI == "BLACKP"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "HISPP"] <- "% Hispanic acres operated"
c$CAT[c$VOI == "HISPP"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "WHITEP"] <- "% white acres operated"
c$CAT[c$VOI == "WHITEP"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "occup"] <- "% primary occupation"
c$CAT[c$VOI == "occup"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "exp"] <- "Years farming"
c$CAT[c$VOI == "exp"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "acres_per_op"] <- "Farm size"
c$CAT[c$VOI == "acres_per_op"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "tenant"] <- "% tenant operators"
c$CAT[c$VOI == "tenant"] <- "B. Farm(er) characteristics"

# Inputs

c$VOIC[c$VOI == "fert"] <- "Fertilizer expense"
c$CAT[c$VOI == "fert"] <- "A. Farm resources"

c$VOIC[c$VOI == "chem"] <- "Chemical expense"
c$CAT[c$VOI == "chem"] <- "A. Farm resources"

c$VOIC[c$VOI == "labor_expense"] <- "Labor expense"
c$CAT[c$VOI == "labor_expense"] <- "A. Farm resources"

c$VOIC[c$VOI == "machinery"] <- "Machinery expense"
c$CAT[c$VOI == "machinery"] <- "A. Farm resources"

c$VOIC[c$VOI == "gvt_prog"] <- "Government receipts"
c$CAT[c$VOI == "gvt_prog"] <- "A. Farm resources"

c$VOIC[c$VOI == "insur_acres"] <- "Crop insurance"
c$CAT[c$VOI == "insur_acres"] <- "A. Farm resources"

# Landscape

c$VOIC[c$VOI == "AREA_MN_AG"] <- "Mean patch area"
c$CAT[c$VOI == "AREA_MN_AG"] <- "C. Landscape"

c$VOIC[c$VOI == "SDI"] <- "Shannon Diversity"
c$CAT[c$VOI == "SDI"] <- "C. Landscape"

c$VOIC[c$VOI == "PNC"] <- "% natural cover"
c$CAT[c$VOI == "PNC"] <- "C. Landscape"

c$VOIC[c$VOI == "ED"] <- "Edge density"
c$CAT[c$VOI == "ED"] <- "C. Landscape"

a <- ggplot(c %>% filter(BS != "Average spot", CAT == "A. Farm resources")) +
  geom_boxplot(aes(x = VOIC, y = value, fill = BS, group = interaction(VOIC, BS)), outlier.size = 0.5,
               outlier.alpha = 0.5) +
  project_theme +
  labs(x = "",
       y = "",
       title = "A. Farm resources",
       fill = "") +
  ylim(c(-3, 3)) +
  coord_flip() +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") +
  theme(text = element_text(size = 16),
        plot.title.position = "plot")

b <- ggplot(c %>% filter(BS != "Average spot", CAT == "B. Farm(er) characteristics")) +
  geom_boxplot(aes(x = VOIC, y = value, fill = BS, group = interaction(VOIC, BS)), outlier.size = 0.5,
               outlier.alpha = 0.5) +
  project_theme +
  labs(x = "",
       y = "",
       title = "B. Farm(er) characteristics",
       fill = "") +
  ylim(c(-3, 3)) +
  coord_flip() +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") +
  theme(text = element_text(size = 16),
        plot.title.position = "plot")

cp <- ggplot(c %>% filter(BS != "Average spot", CAT == "C. Landscape")) +
  geom_boxplot(aes(x = VOIC, y = value, fill = BS, group = interaction(VOIC, BS)), outlier.size = 0.5,
               outlier.alpha = 0.5, varwidth=T) +
  project_theme +
  labs(x = "",
       y = "",
       title = "C. Landscape",
       fill = "") +
  ylim(c(-3, 3)) +
  coord_flip() +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") +
  theme(text = element_text(size = 16),
        plot.title.position = "plot")


ggarrange(ggarrange(a, cp, ncol=1, nrow=2, heights = c(0.9, 0.6), legend = "none"), b, ncol = 2, nrow = 1, common.legend = T, legend = "bottom",
          heights = c(1.1, 0.7), widths = c(.8,1))
ggsave("./figs/allvars.png", width=12, height=10)

##library(cowplot)
#plot_grid(a, b, cp, ncol = 3, align = "v")

# ggplot(c %>% filter(BS != "Average spot")) +
#   geom_boxplot(aes(x = VOIC, y = value, fill = BS, group = interaction(VOIC, BS)), outlier.size = 0.5,
#                outlier.alpha = 0.5) +
#   project_theme +
#   labs(x = "",
#        y = "",
#        title = "") +
#   theme(legend.position = "bottom",
#         axis.text.x = element_text(angle = 45, hjust=1),
#         strip.placement = "above",
#         strip.background = element_rect(fill=NA,colour="black"),
#         panel.spacing=unit(0,"cm"), axis.title.y = element_blank()) +
#   scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") +
#   facet_grid(CAT ~., drop=T, shrink=F, switch="y", space="free", scales="free") +
#   scale_y_continuous(labels = comma) + 
#   # coord_flip() 
#   scale_x_discrete(limits = rev(levels(c$CAT))) +
#   ylim(c(-3, 3)) +
#   theme(legend.text = element_text(size=16),
#         text = element_text(size=16)) +
#   coord_flip() 
# ggsave("./figs/allwa.png", height=14, width=8)


```

***

# Figure 5: Random forests

```{r echo=F, mesasge=F, warning=F, fig.show='hide'}
# length(dfscale$HISPP[is.na(dfscale$HISPP)])/nrow(dfscale)

library(randomForest)
rfData <- dfscale %>% select(c(WAVG, fert, chem, machinery, labor_expense, gvt_prog, insur_acres, 
                                exp, occup, tenant, acres_per_op, 
                                
                                PNC, SDI, ED, AREA_MN_AG
                                )) %>% na.omit()

# build training and testing datasets
set.seed(1)
random_rn <- sample(nrow(rfData), ceiling(nrow(rfData)*.25))
train <- rfData[-random_rn,]
ho <- rfData[random_rn,] 

# tuneRF(x = train[,1:44], y = train$BS_BIN, ntreeTry = 501)
rf_bs <- randomForest(formula = WAVG ~.,
                   data = train,
                   importance = T,
                   mtry = ncol(rfData) - 5)

yhat <- predict(rf_bs, newdata = ho)
mse <- mean((as.numeric(as.character(yhat)) - as.numeric(as.character(ho$WAVG)))^2) # test set MSE

# tuneRF(x = train[,1:44], y = train$BS_BIN, ntreeTry = 501)
rf_bs <- randomForest(formula = WAVG ~.,
                   data = rfData,
                   importance = T,
                   mtry = ncol(rfData) - 5)

imp <- as.data.frame(varImpPlot(rf_bs))
imp$varnames <- rownames(imp) # row names to column
imp$varnames <- as.factor(imp$varnames)
rownames(imp) <- NULL

imp$VOI <- imp$varnames
c <- imp
c$VOIC <- NA
c$CAT <- NA
c$VOIC[c$VOI == "female"] <- "% female operators"
c$CAT[c$VOI == "female"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "BLACKP"] <- "% Black acres operated"
c$CAT[c$VOI == "BLACKP"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "HISPP"] <- "% Hispanic acres operated"
c$CAT[c$VOI == "HISPP"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "WHITEP"] <- "% white acres operated"
c$CAT[c$VOI == "WHITEP"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "occup"] <- "% primary occupation"
c$CAT[c$VOI == "occup"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "exp"] <- "Years farming"
c$CAT[c$VOI == "exp"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "acres_per_op"] <- "Farm size"
c$CAT[c$VOI == "acres_per_op"] <- "B. Farm(er) characteristics"

c$VOIC[c$VOI == "tenant"] <- "% tenant operators"
c$CAT[c$VOI == "tenant"] <- "B. Farm(er) characteristics"

# Inputs

c$VOIC[c$VOI == "fert"] <- "Fertilizer expense"
c$CAT[c$VOI == "fert"] <- "A. Farm resources"

c$VOIC[c$VOI == "chem"] <- "Chemical expense"
c$CAT[c$VOI == "chem"] <- "A. Farm resources"

c$VOIC[c$VOI == "labor_expense"] <- "Labor expense"
c$CAT[c$VOI == "labor_expense"] <- "A. Farm resources"

c$VOIC[c$VOI == "machinery"] <- "Machinery expense"
c$CAT[c$VOI == "machinery"] <- "A. Farm resources"

c$VOIC[c$VOI == "gvt_prog"] <- "Government receipts"
c$CAT[c$VOI == "gvt_prog"] <- "A. Farm resources"

c$VOIC[c$VOI == "insur_acres"] <- "Crop insurance"
c$CAT[c$VOI == "insur_acres"] <- "A. Farm resources"

# Landscape

c$VOIC[c$VOI == "AREA_MN_AG"] <- "Mean patch area"
c$CAT[c$VOI == "AREA_MN_AG"] <- "C. Landscape"

c$VOIC[c$VOI == "SDI"] <- "Shannon Diversity"
c$CAT[c$VOI == "SDI"] <- "C. Landscape"

c$VOIC[c$VOI == "PNC"] <- "% natural cover"
c$CAT[c$VOI == "PNC"] <- "C. Landscape"

c$VOIC[c$VOI == "ED"] <- "Edge density"
c$CAT[c$VOI == "ED"] <- "C. Landscape"

ggplot(c, aes(x = reorder(VOIC, `%IncMSE`), y = `%IncMSE`)) +
  geom_point() +
  geom_segment(aes(x=reorder(VOIC, `%IncMSE`), xend=VOIC,y=0,yend=`%IncMSE`)) +
  ylab("% increase in MSE") +
  xlab("") +
  ggtitle("") +
  coord_flip() +
  project_theme +
  facet_grid(CAT ~ ., drop=T, shrink=F, switch="y", space = "free", scales = "free") +
theme(strip.placement = "outside",
      strip.background = element_rect(fill=NA,colour=NA),
      panel.spacing=unit(0,"cm"), axis.title.y = element_blank(),
      legend.position = "bottom") 

# ggsave("./figs/variable_imp.png", plot = last_plot())

```


# Figure 6: High plains aquifer

Shapefile downloaded [here](https://water.usgs.gov/ogw/gwrp/activities/gspdata/Studies/HighPlains.html).

```{r}
library(sf)
hpa <- st_read("./data/hp_bound2010.shp")
limit <- c(-4, 4)

out <- prep_data("corn")
county_sub <- out[[2]]
df <- prep_attr_data("corn", scale=F)
st_geometry(df) <- NULL
corn_plt <- merge(county_sub, df, by = "GEOID")

st_transform(hpa, st_crs(corn_plt))

corn_sub <- corn_plt %>% filter(STATEFP %in% c("48", "40", "35", "08", "20", "31", "56", "46"))

states <- state %>% filter(STATE_FIPS %in% c("48", "40", "35", "08", "20", "31", "56", "46"))

corn_plot <- ggplot() +
      geom_sf(data = corn_sub, aes(fill = DIFF), color = "transparent") +
    geom_sf(data = states, size = 1, fill = "transparent", color = "gray") +
        geom_sf(data = hpa, size = 1, fill = "transparent") +
    project_theme +
    labs(x = "A. Corn", 
         fill = "") +
    theme(axis.title.x = element_text(size = 15)) +
         scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

out <- prep_data("soy")
county_sub <- out[[2]]
df <- prep_attr_data("soy", scale=F)
st_geometry(df) <- NULL
corn_plt <- merge(county_sub, df, by = "GEOID")

st_transform(hpa, st_crs(corn_plt))

corn_sub <- corn_plt %>% filter(STATEFP %in% c("48", "40", "35", "08", "20", "31", "56", "46"))

soy_plot <- ggplot() +
      geom_sf(data = corn_sub, aes(fill = DIFF), color = "transparent") +
    geom_sf(data = states, size = 1, fill = "transparent", color = "gray") +
        geom_sf(data = hpa, size = 1, fill = "transparent") +
    project_theme +
    labs(x = "C. Soy", 
         fill = "") +
    theme(axis.title.x = element_text(size = 15)) +
         scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

out <- prep_data("hay")
county_sub <- out[[2]]
df <- prep_attr_data("hay", scale=F)
st_geometry(df) <- NULL
corn_plt <- merge(county_sub, df, by = "GEOID")

st_transform(hpa, st_crs(corn_plt))

corn_sub <- corn_plt %>% filter(STATEFP %in% c("48", "40", "35", "08", "20", "31", "56", "46"))

hay_plot <- ggplot() +
      geom_sf(data = corn_sub, aes(fill = DIFF), color = "transparent") +
    geom_sf(data = states, size = 1, fill = "transparent", color = "gray") +
        geom_sf(data = hpa, size = 1, fill = "transparent") +
    project_theme +
    labs(x = "B. Hay", 
         fill = "") +
    theme(axis.title.x = element_text(size = 15)) +
          scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())



library(ggpubr)

ggarrange(corn_plot, hay_plot, soy_plot, ncol = 3, nrow = 1, common.legend = T, legend = "right")

ggsave("./figs/hpa.png", plot = last_plot(), width = 10, height = 6)

```

# Figure 7: Livelihood and PDP

```{r}
dfs %>% group_by(BS) %>% summarize(med = median(gvt_prog, na.rm=T))
plt <- melt(dfscale, id.vars = c("BS"),
            measure.vars = c("crop_sales", "income_farm", "income")) %>%
  filter(BS != "Average")
plt$variable <- as.character(plt$variable)


plt$variable[plt$variable == "income_farm"] <- "Net-farm income ($/operation)"
plt$variable[plt$variable == "income"] <-  "Total farm income ($/operation)"
plt$variable[plt$variable == "crop_sales"] <-  "Crop sales ($/operated acre)"
plt$variable <- as.factor(plt$variable)

lh <- ggplot(plt %>% filter(BS != "Average spot", value <4)) +
  geom_boxplot(aes(x = variable, y = value, fill = BS, group = interaction(variable, BS)), outlier.size = 0.5,
               outlier.alpha = 0.5) +
  project_theme +
  labs(x = "",
       y = "",
       title = "") +
  theme(legend.position = "right") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") +
  #facet_wrap(vars(variable), nrow = length(unique(plt$variable)), scales = "free") +
  scale_y_continuous(labels = comma) + coord_flip() +
  scale_x_discrete(limits = rev(levels(plt$variable))) +
  theme(text = element_text(size = 16))

# ggsave("./figs/livelihood_wa.png", width = 6, height=4)
```

```{r echo=F}
fertp <- as.data.frame(partialPlot(rf_bs, rfData, "fert"))
fertp$VAR = "Fertilizer ($/acre)"                       

chemp <- as.data.frame(partialPlot(rf_bs, rfData, "chem"))
chemp$VAR = "Chemicals ($/acre)" 

labp <- as.data.frame(partialPlot(rf_bs, rfData, "fert"))
labp$VAR = "Fertilizer ($/acre)" 

machp <- as.data.frame(partialPlot(rf_bs, rfData, "machinery"))
machp$VAR = "Machinery ($/acre)" 

gvtp <- as.data.frame(partialPlot(rf_bs, rfData, "gvt_prog"))
gvtp$VAR = "Gvt. receipts ($/acre)" 

insp <- as.data.frame(partialPlot(rf_bs, rfData, "insur_acres"))
insp$VAR = "Crop insurance (% acres))" 

# chemp <- as.data.frame(partialPlot(rf_bs, rfData, "chem"))
dm <- rbind.data.frame(fertp, chemp, labp, gvtp)


input_pp <- ggplot(dm) +
  geom_line(aes(x = x, y = y, color = VAR), size = 1.2) +
  project_theme +
  theme(legend.title = element_blank(), legend.position = "right" ) +
  labs(x = "Standardized values",
       y = "Predicted weighted average") +
  theme(text = element_text(size=12))

#ggsave("./figs/pd_inputs.png", plot = input_pp, width = 6, height = 3)
# https://christophm.github.io/interpretable-ml-book/pdp.html

ggarrange(lh, input_pp, labels = c("A", "B"), ncol=1, nrow=2)
ggsave("./figs/lh_inputs.png", width = 10, height = 8)

```



# SI APPENDIX

# Table 1: Model performance

```{r}
r2 <- rbind.data.frame(dcorn$R2, dsoy$R2, dwwheat$R2, dalfalfa$R2, dhay$R2)
mse <- rbind.data.frame(dcorn$MSE, dsoy$MSE, dwwheat$MSE, dalfalfa$MSE, dhay$MSE)
dic <- rbind.data.frame(dcorn$DIC, dsoy$DIC, dwwheat$DIC, dalfalfa$DIC, dhay$DIC)
cpo <- rbind.data.frame(dcorn$CPO, dsoy$CPO, dwwheat$CPO, dalfalfa$CPO, dhay$CPO)

perf <- cbind.data.frame(clean_crops, r2, mse, dic, cpo)
colnames(perf) <- c("CROP", "R2", "MSE", "DIC", "CPO")

kable(perf %>% select(-c(DIC, CPO)))
```

## Table 2:

### Proportion variance explained

```{r eval=F}
# INLA p. 186
compute_var_exp <- function(crop, model) {
  
  out <- prep_data(crop)[[1]]
  mat.marg <- matrix(NA, nrow = length(unique(out$GEOID)), ncol = 100000)
  m <- model$marginals.random$ID
  
  for (i in 1:length(unique(out$GEOID))) {
    
    u <- m[[length(unique(out$GEOID))+i]]
    mat.marg[i,] <- inla.rmarginal(100000, u)
    
  }
  
  var.u <- apply(mat.marg, 2, var)
  
  var.v <- inla.rmarginal(100000, model$marginals.hyper$`Precision for ID`)
  
  perc.var.u <- mean(var.u/(var.u + var.v))
  return(perc.var.u)
}

pvcorn <- compute_var_exp("corn", mcorn)
pvsoy <- compute_var_exp("soy", msoy)
pvwwheat <- compute_var_exp("wwheat", mwwheat)
pvalfalfa <- compute_var_exp("alfalfa", malfalfa)
pvhay <- compute_var_exp("hay", mhay)
```

### Spatial autocorrelation

See visualizations in `BS_spatial_structure` document.

```{r message=FALSE, warning=FALSE, echo=F, eval=F}
moran_list <- list()
cg_list <- list()

for (i in 1:length(crops)) {
  
  print(crops[[i]])
  crop <- crops[[i]]
  county_sub <- prep_data(crop)[[2]]
  out <- prep_attr_data(crop)
  df <- out[[1]]
  
  county_sp <- as(county_sub, "Spatial")
  county_sp <- merge(county_sp, df, by = "GEOID")
  county_sp <- county_sp[!is.na(county_sp$DIFF),]

  # inverse distance
  queen <- poly2nb(county_sp)
  dists <- nbdists(queen, coordinates(county_sp))
  idw <- lapply(dists, function(x) 1/(x/1000))
  idwlist <- nb2listw(queen, glist = idw, zero.policy = T)
  # gmoran <- moran.test(county_sp$DIFF, listw = idwlist, randomisation = T, zero.policy = T)
  moran <- moran.mc(county_sp$DIFF, idwlist, nsim = 1000, zero.policy = T)
  morandf <- cbind.data.frame(MI = moran$statistic, MP = moran$p.value, CROP = crop)
  rownames(morandf) <- NULL
  moran_list[[i]] <- morandf
  
  # correlogram
  cgm <- sp.correlogram(queen, county_sp$DIFF, order = 25, method = "I", zero.policy = T, randomisation=T)
  cg_list[[i]] <- cgm

}

saveRDS(cg_list, "./out/correlograms.RDS")
saveRDS(moran_list, "./out/morans.RDS")
```

# Figure 1: USDA Farm resource regions

```{r}
ggplot() +
      geom_sf(data = frr_shp, size = 1, aes(fill = as.factor(NAME))) +
  geom_sf(data = state, color = "#A9A9A9", fill = "gray", alpha = 0.5) +

  scale_fill_brewer(type = "qual", palette = "Accent") +
  project_theme +
  labs(fill = "",
       title = "",
       subtitle = "") +
  labs(fill = "",
       title = "",
       subtitle = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank()) 
ggsave("./figs/FRR.png")
```


# Figure 2: Regional plots


```{r}
limit <- c(-1, 1)

total_area_effects <- function(model_run, data, level, county_sub) {
  
  r <- model_run
  n <- length(unique(data[,level]))
  
  # extract area-specific residuals, random effects, Epsilon = u + v
  asr <- r$summary.random[[level]][1:n, c(1, 2, 3)]  # ID, mean, sd
  county_sp <- as(county_sub, "Spatial")
  
  if (level == "FRR") {
    asr <- r$summary.random[[level]][1:(n + 2), c(1, 2, 3)]  # ID, mean, sd
    colnames(asr) <- c("FRR", "mean", "SD")
  } 
  
  map_asr <- merge(county_sp, asr, by = level, all=T)
  map_asr <- merge(county, map_asr, by = 'GEOID', all = T)
  
  limit <- c(-1, 1)
  se <- ggplot() +
    geom_sf(data = map_asr, color = "transparent", size = 0.05, aes(fill = mean)) +
    geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
    geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
    project_theme +
    labs(fill = "") +
    scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                         limit = limit) +
    labs(fill = "",
         title = "",
         subtitle = "") +
    theme(text = element_text(size = 20)) +
    theme(axis.text.x = element_blank(),
          axis.text.y = element_blank(),
          axis.ticks = element_blank(),
          rect = element_blank())
  
  
  return(se)
  
}



make_reg_plot <- function(crop, modr, title) {
  
  out <- prep_data(crop, scale=T)
  null <- out[[1]]
  county_sub <- out[[2]]
  modlrr <- total_area_effects(modr, null, level="FRR", county_sub) + ggtitle(title)
  # plt <- modlrr + ggtitle("Regional IID effects") +
  #     scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", 
  #                          midpoint = 0, na.value = "#D3D3D3", 
  #                         limit = limit) +
  #   labs(fill = "",
  #        title = title,
  #        subtitle = "") + theme(text = element_text(size = 30)) +
  #   theme(axis.text.x = element_blank(),
  #         axis.text.y = element_blank(),
  #         axis.ticks = element_blank(),
  #         rect = element_blank())
  return(modlrr)

}

cp <- make_reg_plot("corn", mcorn, "C. Corn")
sp <- make_reg_plot("soy", msoy, "D. Soy")
wp <- make_reg_plot("wwheat", mwwheat, "E. Winter wheat")
ap <- make_reg_plot("alfalfa", malfalfa, "A. Alfalfa")
hp <- make_reg_plot("hay", mhay, "B. Hay")

# ACROSS crop bright spots
dfs <- df2 %>%
  group_by(FRR.x) %>%
  summarize(N = sum(PROP_CROP, na.rm=T),
    #N = length(DIFF[!is.na(DIFF)]),
            TOTAL = sum(WD, na.rm=T)) %>%
  mutate(WAVG = TOTAL/N)

wd <- ggplot() +
  geom_sf(data = dfs, color = "transparent", size = 0.05, aes(fill = WAVG)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
  scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  labs(fill = "",
       title = "",
       subtitle = "") +
  labs(fill = "",
       title = "F. Weighted average",
       subtitle = "") +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())  +   theme(text = element_text(size = 20)) 


ggarrange(ap, hp, cp, sp, wp, wd, ncol = 2, nrow = 3, common.legend = T, legend = "right") 
ggsave("./figs/reg_plot_crop.png", width = 15 , height= 15)

# http://www.sthda.com/english/wiki/wiki.php?id_contents=7930#grid.arrange-create-and-arrange-multiple-plots
#library(gridExtra)
##grid.arrange(ap, hp, cp, sp, wp, wd, lout, ncol = 2, nrow = 4)
#ggsave("./figs/reg_plot_crop.png", width = 30 , height= 30)

```

## Figure 3: Inputs by crop

```{r echo=F}
df$BS[df$BS == "Bright spot"] <- "> 1 sd"
df$BS[df$BS == "Dark spot"] <- "< 1 sd"

fert <- ggplot(df %>% filter(BS != "Average spot", fert < 125)) +
  geom_boxplot(aes(x = CCROP, y = fert, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Fertilizer ($/acre)") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

# chemical expense $/operated acre
chem <- ggplot(df %>% filter(BS != "Average spot", chem < 125)) +
  geom_boxplot(aes(x = CCROP, y = chem, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Chemicals ($/acre)") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

# machinery assests, $/operated acre
mach <- ggplot(df %>% filter(BS != "Average spot", machinery < 1250)) +
  geom_boxplot(aes(x = CCROP, y = machinery, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Machinery ($/acre)") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

labor <- ggplot(df %>% filter(BS != "Average spot", labor_expense < 200)) +
  geom_boxplot(aes(x = CCROP, y = labor_expense, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Labor ($/acre)") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

ins <- ggplot(df %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = insur_acres, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "% acres with federal crop insurance") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

gvt <- ggplot(df %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = gvt_prog, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Gvt. receipts ($/acre)") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

ggarrange(fert, chem, mach, labor, gvt, ins, nrow = 3, ncol = 2, common.legend = T, legend = "bottom",
          labels = c("A", "B", "C", "D", "E", "F"))
ggsave("./figs/inputs.png", width = 10, height = 10)

```


## Characteristics

```{r}
occup <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = occup, fill = BS, group = interaction(CCROP, BS)), size=0.3, outlier.alpha = 0.3, outlier.size = 0.5) +
  project_theme +
  labs(x = "",
       y = "% primary occupation as farming") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "")

exp <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = exp, fill = BS, group = interaction(CCROP, BS)), size=0.3, outlier.alpha = 0.3, outlier.size = 0.5) +
  project_theme +
  labs(x = "",
       y = "Years farming") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

tenant <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = tenant, fill = BS, group = interaction(CCROP, BS)), size=0.3, outlier.alpha = 0.3, outlier.size = 0.5) +
  project_theme +
  labs(x = "",
       y = "% acres operated by tenants") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

apo <- ggplot(df2 %>% filter(BS != "Average spot", acres_per_op < 2000)) +
  geom_boxplot(aes(x = CCROP, y = acres_per_op, fill = BS, group = interaction(CCROP, BS)), size=0.3, outlier.alpha = 0.3, outlier.size = 0.5) +
  project_theme +
  labs(x = "",
       y = "Median acres per operation") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 


## Race by crop
# SI
black <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = BLACKP, fill = BS, group = interaction(CCROP, BS)), size=0.3, outlier.alpha = 0.3, outlier.size = 0.5) +
  project_theme +
  labs(x = "",
       y = "% Black acres operated") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 
 # ylim(c(0,100)) 

hisp <- ggplot(df2 %>% filter(BS != "Average spot", HISPP < 10)) +
  geom_boxplot(aes(x = CCROP, y = HISPP, fill = BS, group = interaction(CCROP, BS)), size=0.3, outlier.alpha = 0.3,
               outlier.size = 0.5) +
  project_theme +
  labs(x = "",
       y = "% Hispanic acres operated") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 
#  ylim(c(0,100))

white <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = WHITEP, fill = BS, group = interaction(CCROP, BS)), size=0.3, outlier.alpha = 0.3, outlier.size = 0.5) +
  project_theme +
  labs(x = "",
       y = "% White acres operated") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 
#  ylim(c(0,100)) 

female <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = female, fill = BS, group = interaction(CCROP, BS)), size=0.3, outlier.alpha = 0.3, outlier.size = 0.5) +
  project_theme +
  labs(x = "",
       y = "% female acres operated") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

ggarrange(occup, exp, tenant, apo, black, hisp, white, female, ncol=2, nrow=4, common.legend=T, legend = "bottom",
          labels = c("A", "B", "C", "D", "E", "F", "G", "H"))
ggsave("./figs/characteristics.png", height=12, width=8)
```


## Landscape by crop

```{r}
# all indices are for AG

sdi <- ggplot(df %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = SDI, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Shannon Diversity Index") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

ed <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = ED, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Edge density") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

ama <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = AREA_MN_AG, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Mean patch area") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

pnc <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = PNC, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Percent natural cover") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

ggarrange(ama, ed, sdi, pnc, ncol = 2, nrow = 2, common.legend = T, legend = "bottom", labels = c("A", "B", "C", "D"))
ggsave("./figs/landscape.png", width=7, height=7)


```


## Livelihood by crop

```{r}
sdi <- ggplot(df %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = income_farm, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Net income ($/operation)") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

ed <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = income, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Total farm income ($/operation)") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

pnc <- ggplot(df2 %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = CCROP, y = crop_sales, fill = BS, group = interaction(CCROP, BS))) +
  project_theme +
  labs(x = "",
       y = "Crop sales ($/operated acre)") +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") 

ggarrange(ed, sdi, pnc, ncol = 2, nrow = 2, common.legend = T, legend = "bottom", labels = c("A", "B", "C"))
ggsave("./figs/livelihoods.png", width=7, height=7)


```


## Race maps


```{r eval=F}
limit = c(0, 100)
# df2$BLACKP[df2$BLACKP > 20] <- 20
blk_map <- ggplot(df2 %>% filter(CROP == "corn")) +
  geom_sf(color = "transparent", size = 0.05, aes(fill = BLACKP)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
  scale_fill_viridis(na.value = "#D3D3D3", limit = limit) +
  #scale_fill_gradient(low = "#8b0000", high = "#000080", na.value = "#D3D3D3" ) +
  labs(fill = "",
       title = "% Black operators",
       subtitle = "") + theme(text = element_text(size = 15), legend.text = element_text(size = 20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())

hisp_map <- ggplot(df2 %>% filter(CROP == "corn")) +
  geom_sf(color = "transparent", size = 0.05, aes(fill = HISPP)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
  scale_fill_viridis(na.value = "#D3D3D3", limit=limit) +
  #scale_fill_gradient(low = "#8b0000", high = "#000080", na.value = "#D3D3D3" ) +
  labs(fill = "",
       title = "% Hispanic operators",
       subtitle = "") + theme(text = element_text(size = 15), legend.text = element_text(size = 20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank())


white_map <- ggplot(df2 %>% filter(CROP == "corn")) +
  geom_sf(color = "transparent", size = 0.05, aes(fill = WHITEP)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
  scale_fill_viridis(na.value = "#D3D3D3", limit = limit) +
  #scale_fill_gradient(low = "#8b0000", high = "#000080", na.value = "#D3D3D3" ) +
  labs(fill = "",
       title = "% white operators",
       subtitle = "") + theme(text = element_text(size = 15), legend.text = element_text(size = 20)) +
  theme(axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank(),
        rect = element_blank()) 

ggarrange(blk_map, hisp_map, white_map, ncol = 1, nrow = 3, labels = c("A", "B", "C"))

ggsave("./figs/racemaps.png", height=12, width=8)
```


## Partial dependence plots

The y-axis of a partial dependence plot for regression represents the marginal impact of the independent variable to the dependent variable. E.g. if the line is at 0, then for that value of the independent variable, there is 0 impact to the dependent variable. It does not represent the predicted value or relative impact of this variable to other variables. However, you can deduce relative impact by plotting multiple partial dependence plots side by side for multiple independent variables and share the y-axis.

s says that the plots show the relative logit contribution of the variable on the class probability from the perspective of the model. In other words negative values (in the y-axis) mean that the positive class is less likely for that value of the independent variable (x-axis) according to the model. Similarly positive values mean that the positive class is more likely for that value of the independent variable according to the model. Clearly, zero implies no average impact on class probability according to the model.



```{r echo=F, mesasge=F, warning=F, fig.show='hide'}
library(randomForest)
rfData <- dfs %>% select(c(WAVG, fert, chem, machinery, labor_expense, gvt_prog, insur_acres, 
                                exp, occup, tenant, acres_per_op, 
                                WHITEP, BLACKP, HISPP, female,
                                PNC, SDI, ED, AREA_MN_AG
                                )) %>% na.omit()

# build training and testing datasets
set.seed(1)
random_rn <- sample(nrow(rfData), ceiling(nrow(rfData)*.25))
train <- rfData[-random_rn,]
ho <- rfData[random_rn,] 

# tuneRF(x = train[,1:44], y = train$BS_BIN, ntreeTry = 501)
rf_bs <- randomForest(formula = WAVG ~.,
                   data = train,
                   importance = T,
                   mtry = ncol(rfData) - 5)

yhat <- predict(rf_bs, newdata = ho)
mse <- mean((as.numeric(as.character(yhat)) - as.numeric(as.character(ho$WAVG)))^2) # test set MSE

# tuneRF(x = train[,1:44], y = train$BS_BIN, ntreeTry = 501)
rf_bs <- randomForest(formula = WAVG ~.,
                   data = rfData,
                   importance = T,
                   mtry = ncol(rfData) - 5)

imp <- as.data.frame(varImpPlot(rf_bs))
imp$varnames <- rownames(imp) # row names to column
imp$varnames <- as.factor(imp$varnames)
rownames(imp) <- NULL

imp$VOI <- imp$varnames
c <- imp
c$VOIC <- NA
c$CAT <- NA
c$VOIC[c$VOI == "female"] <- "% female operators"
c$CAT[c$VOI == "female"] <- "Farm(er) characteristics"

c$VOIC[c$VOI == "BLACKP"] <- "% acres operated by black farmers"
c$CAT[c$VOI == "BLACKP"] <- "Farm(er) characteristics"

c$VOIC[c$VOI == "HISPP"] <- "% acres operated by Hispanic farmers"
c$CAT[c$VOI == "HISPP"] <- "Farm(er) characteristics"

c$VOIC[c$VOI == "WHITEP"] <- "% acres operated by white farmers"
c$CAT[c$VOI == "WHITEP"] <- "Farm(er) characteristics"

c$VOIC[c$VOI == "occup"] <- "% primary occupation"
c$CAT[c$VOI == "occup"] <- "Farm(er) characteristics"

c$VOIC[c$VOI == "exp"] <- "Average years farming"
c$CAT[c$VOI == "exp"] <- "Farm(er) characteristics"

c$VOIC[c$VOI == "acres_per_op"] <- "Acres per operation"
c$CAT[c$VOI == "acres_per_op"] <- "Farm(er) characteristics"

c$VOIC[c$VOI == "tenant"] <- "% tenant operators"
c$CAT[c$VOI == "tenant"] <- "Farm(er) characteristics"

# Inputs

c$VOIC[c$VOI == "fert"] <- "Fertilizer expense ($/acre)"
c$CAT[c$VOI == "fert"] <- "Decision-making"

c$VOIC[c$VOI == "chem"] <- "Chemical expense ($/acre)"
c$CAT[c$VOI == "chem"] <- "Decision-making"

c$VOIC[c$VOI == "labor_expense"] <- "Labor expense ($/acre)"
c$CAT[c$VOI == "labor_expense"] <- "Decision-making"

c$VOIC[c$VOI == "machinery"] <- "Machinery ($/acre)"
c$CAT[c$VOI == "machinery"] <- "Decision-making"

c$VOIC[c$VOI == "gvt_prog"] <- "Government receipts ($/acre)"
c$CAT[c$VOI == "gvt_prog"] <- "Decision-making"

c$VOIC[c$VOI == "insur_acres"] <- "Crop insurance (% acres)"
c$CAT[c$VOI == "insur_acres"] <- "Decision-making"

# Landscape

c$VOIC[c$VOI == "AREA_MN_AG"] <- "Mean patch area"
c$CAT[c$VOI == "AREA_MN_AG"] <- "Landscape"

c$VOIC[c$VOI == "SDI"] <- "Shannon Diversity"
c$CAT[c$VOI == "SDI"] <- "Landscape"

c$VOIC[c$VOI == "PNC"] <- "% natural cover"
c$CAT[c$VOI == "PNC"] <- "Landscape"

c$VOIC[c$VOI == "ED"] <- "Edge density"
c$CAT[c$VOI == "ED"] <- "Landscape"

ggplot(c, aes(x = reorder(VOIC, `%IncMSE`), y = `%IncMSE`)) +
  geom_point() +
  geom_segment(aes(x=reorder(VOIC, `%IncMSE`), xend=VOIC,y=0,yend=`%IncMSE`)) +
  ylab("% increase in MSE") +
  xlab("") +
  ggtitle("") +
  coord_flip() +
  project_theme +
  facet_grid(CAT ~ ., drop=T, shrink=F, switch="y", space = "free", scales = "free") +
theme(strip.placement = "outside",
      strip.background = element_rect(fill=NA,colour=NA),
      panel.spacing=unit(0,"cm"), axis.title.y = element_blank(),
      legend.position = "bottom") 

# ggsave("./figs/variable_imp.png", plot = last_plot())

```


```{r echo=F}
# use unscaled for clarity
fertp <- as.data.frame(partialPlot(rf_bs, rfData, "BLACKP")) ###
fertp$VAR = "% acres operated by black farmers"                       

chemp <- as.data.frame(partialPlot(rf_bs, rfData, "WHITEP"))
chemp$VAR = "% acres operated by white farmers" 

labp <- as.data.frame(partialPlot(rf_bs, rfData, "HISPP"))
labp$VAR = "% acres operated by Hispanic farmers" 

machp <- as.data.frame(partialPlot(rf_bs, rfData, "female"))
machp$VAR = "% female operators" 

rp <- rbind.data.frame(fertp, chemp, labp, machp)

race_pp <- ggplot(rp) +
  geom_line(aes(x = x, y = y, color = VAR), size = 1.2) +
  project_theme +
  theme(legend.title = element_blank(), legend.position = "right" ) +
  labs(x = "% operated land",
       y = "Predicted weighted average")
race_pp


ggsave("./figs/pd_race.png", width = 4, height = 3)
```

```{r echo=F}
fertp <- as.data.frame(partialPlot(rf_bs, rfData, "acres_per_op")) ###
fertp$VAR = "Average farm size"                       

chemp <- as.data.frame(partialPlot(rf_bs, rfData, "exp"))
chemp$VAR = "Average years farming" 

labp <- as.data.frame(partialPlot(rf_bs, rfData, "occup"))
labp$VAR = "% primary occupation" 

machp <- as.data.frame(partialPlot(rf_bs, rfData, "tenant"))
machp$VAR = "% tenants" 

rp <- rbind.data.frame(fertp, chemp, labp, machp)

char_pp <- ggplot(rp) +
  geom_line(aes(x = x, y = y, color = VAR), size = 1.2) +
  project_theme +
  theme(legend.title = element_blank(), legend.position = "right" ) +
  labs(x = "Standardized values",
       y = "Predicted weighted average")

ggsave("./figs/pd_char.png", plot = char_pp, width = 4, height = 3)
```

```{r echo=F}
fertp <- as.data.frame(partialPlot(rf_bs, rfData, "SDI")) ###
fertp$VAR = "Shannon Diversity of agriculture"                       

chemp <- as.data.frame(partialPlot(rf_bs, rfData, "PNC"))
chemp$VAR = "% natural cover" 

labp <- as.data.frame(partialPlot(rf_bs, rfData, "ED"))
labp$VAR = "Edge density" 

machp <- as.data.frame(partialPlot(rf_bs, rfData, "AREA_MN_AG"))
machp$VAR = "Mean patch area" 

lp <- rbind.data.frame(fertp, labp, machp)

l_pp <- ggplot(lp) +
  geom_line(aes(x = x, y = y, color = VAR), size = 1.2) +
  project_theme +
  theme(legend.title = element_blank(), legend.position = "right" ) +
  labs(x = "Standardized values",
       y = "Predicted weighted average")

ggsave("./figs/pd_ls.png", plot = l_pp, width = 6, height = 3)

ggarrange(char_pp, l_pp, labels = c("A", "B"), ncol = 1, nrow=2)
ggsave("./figs/pdp.png", width=12, height=10)
```

```{r echo=F}
# top 5 predictors: fert, gvt, black, nat, machinwet
fertp <- as.data.frame(partialPlot(rf_bs, rfData, "acres_per_op")) ###
fertp$VAR = "Average farm size"                       

chemp <- as.data.frame(partialPlot(rf_bs, rfData, "PNC"))
chemp$VAR = "% natural cover" 

labp <- as.data.frame(partialPlot(rf_bs, rfData, "fert"))
labp$VAR = "Fertilizer ($/acre)" 

machp <- as.data.frame(partialPlot(rf_bs, rfData, "BLACKP"))
machp$VAR = "% acres operated by black farmers" 

blp <- as.data.frame(partialPlot(rf_bs, rfData, "machinery"))
blp$VAR = "Machinery ($/acre)" 

top5 <- rbind.data.frame(fertp, chemp, labp, machp, blp, gvtp)

top5_pp <- ggplot(top5) +
  geom_line(aes(x = x, y = y, color = VAR), size = 1.2) +
  project_theme +
  theme(legend.title = element_blank(), legend.position = "right" ) +
  labs(x = "Standardized values",
       y = "Predicted weighted average")

ggsave("./figs/pd_top5.png", width = 4, height = 3)
```


# Archive plots

## Weighted average bw plots

```{r}
# Archive plots
plt$variable[plt$variable == "fert"] <- "(a) Fertilizer ($/acre)"
plt$variable[plt$variable == "chem"] <-  "(b) Chemicals ($/acre)"
plt$variable[plt$variable == "machinery"] <-  "(c) Machinery ($/acre)"
plt$variable[plt$variable == "labor_expense"] <-  "(d) Labor ($/acre)"
plt$variable[plt$variable == "gvt_prog"] <-  "(e) Gvt. receipts ($/acre)"
plt$variable[plt$variable == "insur_acres"] <-  "(f) Crop insurance (% operated acres)"
plt$variable <- as.factor(plt$variable)

input_bw <- ggplot(plt %>% filter(BS != "Average spot", value <4)) +
  geom_boxplot(aes(x = variable, y = value, fill = BS, group = interaction(variable, BS)), outlier.size = 0.5,
               outlier.alpha = 0.5) +
  project_theme +
  labs(x = "",
       y = "",
       title = "") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") +
  #facet_wrap(vars(variable), nrow = length(unique(plt$variable)), scales = "free") +
  scale_y_continuous(labels = comma) + coord_flip() +
  scale_x_discrete(limits = rev(levels(plt$variable)))

# ggsave("./figs/inputs_wa.png", width = 6, height=6)
```

```{r}
# dfs %>% group_by(BS) %>% summarize(med = median(BLACKP, na.rm=T))

plt <- melt(dfscale, id.vars = c("BS"),
            measure.vars = c("exp", "tenant", "occup", "acres_per_op")) %>%
  filter(BS != "Average")
plt$variable <- as.character(plt$variable)

plt$variable[plt$variable == "exp"] <- "(a) Years farming experience"
plt$variable[plt$variable == "occup"] <-  "(b) % farming as primary occupation"
plt$variable[plt$variable == "tenant"] <-  "(c) % tenants"
plt$variable[plt$variable == "acres_per_op"] <-  "(d) Average farm size"
plt$variable <- as.factor(plt$variable)

char_bw <- ggplot(plt %>% filter(BS != "Average spot", value <4)) +
  geom_boxplot(aes(x = variable, y = value, fill = BS, group = interaction(variable, BS)), outlier.size = 0.5,
               outlier.alpha = 0.5) +
  project_theme +
  labs(x = "",
       y = "",
       title = "") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") +
  #facet_wrap(vars(variable), nrow = length(unique(plt$variable)), scales = "free") +
  scale_y_continuous(labels = comma) + coord_flip() +
  scale_x_discrete(limits = rev(levels(plt$variable)))

# ggsave("./figs/character_wa.png", width = 6, height=4)
```

```{r eval=F}
race <- read.csv("./data/race.csv")
#unique(race$Data.Item)
race$Value[race$Value == "999"] <- NA
race$Value <- as.numeric(gsub(",", "", race$Value))  

black <- race %>%
  filter(Data.Item == "OPERATORS, PRINCIPAL, BLACK OR AFRICAN AMERICAN - ACRES OPERATED") %>%
  filter(!is.na(County.ANSI)) %>%
  mutate(STATE = str_pad(as.character(State.ANSI), 2, side = "left", pad = "0"),
         COUNTY = str_pad(as.character(County.ANSI), 3, side = "left", pad = "0"),
         GEOID = as.factor(paste0(STATE,COUNTY)),
         BLACK = Value,
         YEAR = Year) %>%
  select(YEAR, GEOID, BLACK) 

h <- race %>%
  filter(Data.Item == "OPERATORS, PRINCIPAL, HISPANIC - ACRES OPERATED"  ) %>%
  filter(!is.na(County.ANSI)) %>%
  mutate(STATE = str_pad(as.character(State.ANSI), 2, side = "left", pad = "0"),
         COUNTY = str_pad(as.character(County.ANSI), 3, side = "left", pad = "0"),
         GEOID = as.factor(paste0(STATE,COUNTY)),
         HISP = Value,
         YEAR = Year) %>%
  select(YEAR, GEOID, HISP) 

w <- race %>%
  filter(Data.Item ==  "OPERATORS, PRINCIPAL, WHITE - ACRES OPERATED") %>%
  filter(!is.na(County.ANSI)) %>%
  mutate(STATE = str_pad(as.character(State.ANSI), 2, side = "left", pad = "0"),
         COUNTY = str_pad(as.character(County.ANSI), 3, side = "left", pad = "0"),
         GEOID = as.factor(paste0(STATE,COUNTY)),
         WHITE = Value,
         YEAR = Year) %>%
  select(YEAR, GEOID, WHITE) 

out <- merge(black, h, by = c("GEOID", "YEAR"), all=T)
out <- merge(out, w, by = c("GEOID", "YEAR"), all=T)

# divide by total operated acres to compute % operated acres
ao <- read.csv("./data/acres.csv")
ao$Value[ao$Value == "999"] <- NA
ao$Value <- as.numeric(gsub(",", "", ao$Value))  
ao <- ao %>%
  filter(!is.na(County.ANSI)) %>%
  mutate(STATE = str_pad(as.character(State.ANSI), 2, side = "left", pad = "0"),
         COUNTY = str_pad(as.character(County.ANSI), 3, side = "left", pad = "0"),
         GEOID = as.factor(paste0(STATE,COUNTY)),
         ACRES_AG = Value,
         YEAR = Year) %>%
  select(YEAR, GEOID, ACRES_AG) 

out <- merge(out, ao, by = c("GEOID", "YEAR"), all=T)

# need to think about size of entire county?  Will bias results in really big counties... but WHITE consistently more than total ag acrea.

# divide by total operated acres to compute % operated acres
ao <- read.csv("./data/organic.csv")
ao$Value[ao$Value == "999"] <- NA
ao$Value <- as.numeric(gsub(",", "", ao$Value))  
ao <- ao %>%
  filter(!is.na(County.ANSI)) %>%
  mutate(STATE = str_pad(as.character(State.ANSI), 2, side = "left", pad = "0"),
         COUNTY = str_pad(as.character(County.ANSI), 3, side = "left", pad = "0"),
         GEOID = as.factor(paste0(STATE,COUNTY)),
         ORGANIC_ACRES = Value,
         YEAR = Year) %>%
  select(YEAR, GEOID, ORGANIC_ACRES) 

out <- merge(out, ao, by = c("GEOID", "YEAR"), all = T)

# load total ag area
# ao <- read.csv("C:/Users/eburchf/OneDrive - Emory University/WF/Irrigation/data/ag_land_acres.csv")
# ao$Value[ao$Value == "999"] <- NA
# ao$Value <- as.numeric(gsub(",", "", ao$Value))  
# ao <- ao %>%
#   filter(!is.na(County.ANSI)) %>%
#   mutate(STATE = str_pad(as.character(State.ANSI), 2, side = "left", pad = "0"),
#          COUNTY = str_pad(as.character(County.ANSI), 3, side = "left", pad = "0"),
#          GEOID = as.factor(paste0(STATE,COUNTY)),
#          CL_ACRES = Value,
#          YEAR = Year) %>%
#   select(YEAR, GEOID, CL_ACRES) 

# from BS
std <- readRDS("./data/standardize.RDS")
out <- merge(out, std, by = c("GEOID", "YEAR"), all = T)


# out <- merge(out, ao, by = c("GEOID", "YEAR"), all = T) %>% filter(YEAR %in% c("2017", "2012"))
out$BLACKP <- out$BLACK/out$OPERATED_ACRES * 100
out$HISPP <- out$HISP/out$OPERATED_ACRES * 100
out$WHITEP <- out$WHITE/out$OPERATED_ACRES * 100
out$BIOP <- out$ORGANIC_ACRES/out$OPERATED_ACRES * 100

saveRDS(out, "./out/race.RDS")
```

```{r}

plt <- melt(dfs, id.vars = c("BS"),
            measure.vars = c("BLACKP", "WHITEP", "HISPP", "female")) %>%
  filter(BS != "Average")
plt$variable <- as.character(plt$variable)

plt$variable[plt$variable == "BLACKP"] <- "(a) Black farmers"
plt$variable[plt$variable == "HISPP"] <-  "(b) Hispanic farmers"
plt$variable[plt$variable == "female"] <-  "(c) Female farmers"
plt$variable[plt$variable == "WHITEP"] <-  "(d) White farmers"

race_bw <- ggplot(plt %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = variable, y = value, fill = BS, group = interaction(variable, BS)), outlier.size = 0.5,
               outlier.alpha = 0.5) +
  project_theme +
  labs(x = "",
       y = "% operated acres",
       title = "") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") +
  #facet_wrap(vars(variable), nrow = length(unique(plt$variable)), scales = "free") +
  scale_y_continuous(labels = comma) + coord_flip() +
  scale_x_discrete(limits = rev(levels(plt$variable)))

# ggsave("./figs/race_wa.png", width = 6, height=4)

```


```{r}
plt <- melt(dfscale, id.vars = c("BS"),
            measure.vars = c("AREA_MN_AG", "SDI", "PNC", "ED")) %>%
  filter(BS != "Average")
plt$variable <- as.character(plt$variable)

plt$variable[plt$variable == "AREA_MN_AG"] <- "(a) Mean patch size"
plt$variable[plt$variable == "ED"] <-  "(b) Edge density"
plt$variable[plt$variable == "SDI"] <-  "(c) Shannon Diversity of agriculture"
plt$variable[plt$variable == "PNC"] <-  "(d) Percent natural cover"

ggplot(plt %>% filter(BS != "Average spot")) +
  geom_boxplot(aes(x = variable, y = value, fill = BS, group = interaction(variable, BS)), outlier.size = 0.5,
               outlier.alpha = 0.5) +
  project_theme +
  labs(x = "",
       y = "",
       title = "") +
  theme(legend.position = "right",
        axis.text.x = element_text(angle = 45, hjust=1)) +
  scale_fill_manual(values = c("#317EBE", "#4C9A2A"), name = "") +
  #facet_wrap(vars(variable), nrow = length(unique(plt$variable)), scales = "free") +
  scale_y_continuous(labels = comma) + coord_flip() +
  scale_x_discrete(limits = rev(levels(plt$variable)))

ggsave("./figs/landscape_wa.png", width = 6, height=4)
```


## Weighted difference

```{r}
# ACROSS crop bright spots
dfs <- dfus %>%
  group_by(GEOID) %>%
  summarize(N = sum(PROP_CROP, na.rm=T),
    #N = length(DIFF[!is.na(DIFF)]),
            TOTAL = sum(WD, na.rm=T)) %>%
  mutate(WAVG = TOTAL/N)
dfa <-  dfus %>%
    select(-c(BS, geometry, WD, PROP_CROP, CROP, CCROP, Z, DIFF, STATEFP, FRR.x, FRR.y)) %>% # remove null model attr
    group_by(GEOID) %>% 
    summarize_all(funs(mean), na.rm=T)
st_geometry(dfs) <- NULL
st_geometry(dfa) <- NULL
dfs <- merge(county, dfs, by = "GEOID", all=T)
dfs <- merge(dfs, dfa, by = "GEOID", all = T)
dfs$BS <- ifelse(dfs$WAVG > mean(dfs$WAVG, na.rm=T) + 1.5*sd(dfs$WAVG, na.rm=T), "Bright spot", NA)
dfs$BS[dfs$WAVG < mean(dfs$WAVG, na.rm=T) - 1.5*sd(dfs$WAVG, na.rm=T)] <- "Dark spot"
dfs$BS[is.na(dfs$BS)] <- "Average spot"

limit <- max(abs(dfs$WAVG), na.rm=T) * c(-1, 1)

ggplot() +
  geom_sf(data = dfs, color = "transparent", size = 0.05, aes(fill = WAVG)) +
  geom_sf(data = state, color = "#A9A9A9", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = NA, alpha = 0.5) +
  project_theme +
  scale_fill_gradient2(low = "#654321", mid = "#f5f5f5", high = "#000080", midpoint = 0, na.value = "#D3D3D3", 
                       limit = limit) +
  labs(fill = "",
       title = "",
       subtitle = "")

ggsave("./figs/weighted_diff_avg.png", plot = last_plot())
```

