---
title: "BS Model Comparison"
author: "Dr. Burchfield"
output: html_document
---

## Results {.tabset .tabset-fade}

### Model speficiation

```{r eval=F}
  out <- prep_data(crop)
  null <- out[[1]]
  county_sub <- out[[2]]
  hadj <- build_adj_matrix(county_sub)
  
  null <- null %>% mutate_at(c("GDD", "TP"), ~inla.group(., n = 20, method = "quantile"))
  u <- sd(null$YIELD, na.rm = T)  
  
  apar <- 0.5 
  lmod <- lm(YIELD ~ GDD + TP  + T_CEC_SOIL + T_OC + T_PH_H2O + T_ESP + 
               PERC_IRR + 
               factor(YEAR), data=  null)
  bpar <- apar*var(residuals(lmod))
  lgprior <- list(prec = list(prior = "loggamma", param = c(apar, bpar))) 
  
  formula <- YIELD ~ 1 +  
    f(ID, model="bym2", 
      graph=hadj,
      scale.model=TRUE,
      constr = TRUE) +
    f(FRR, model = "iid", hyper = lgprior) +
    f(GDD, model = "rw2", 
      scale.model = T,
      hyper = list(theta = list(prior = "pc.prec", 
                                param = c(u, 0.01)))) +
    f(TP, model = "rw2", 
      scale.model = T,
      hyper = list(theta = list(prior = "pc.prec", 
                                param = c(u, 0.01)))) +
    PERC_IRR +
    T_CEC_SOIL + T_OC + T_PH_H2O + T_ESP + 
    factor(YEAR)
  
```


```{r echo=F, message=F, warning=F}
library(tidyverse)
source("BS_load.R")
source("BS_func.R")
mcorn <- readRDS("./out/models/mf_corn.RDS")
dcorn <- readRDS("./out/models/df_corn.RDS")

msoy <- readRDS("./out/models/mf_soy.RDS")
dsoy <- readRDS("./out/models/df_soy.RDS")

mwwheat <- readRDS("./out/models/mf_wwheat.RDS")
dwwheat <- readRDS("./out/models/df_wwheat.RDS")

malfalfa <- readRDS("./out/models/mf_alfalfa.RDS")
dalfalfa <- readRDS("./out/models/df_alfalfa.RDS")

mhay <- readRDS("./out/models/mf_hay.RDS")
dhay <- readRDS("./out/models/df_hay.RDS")

msorghum <- readRDS("./out/models/mf_sorghum.RDS")
dsorghum <- readRDS("./out/models/df_sorghum.RDS")

mcotton <- readRDS("./out/models/mf_cotton.RDS")
dcotton <- readRDS("./out/models/df_cotton.RDS")

mlist <- list(mcorn, msoy, mwwheat, malfalfa, msorghum, mcotton, mhay)

# crop colors
#https://www.nature.com/articles/nmeth.1618
wong_colors <- c("#E69F00", "#55B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

library(RColorBrewer)

crops <- as.factor(c("corn", "soy", "wwheat", "alfalfa", "sorghum", "cotton", "hay"))
clean_crops <- as.factor(c("Corn", "Soy", "Winter wheat", "Alfalfa", "Sorghum", "Cotton", "Hay"))
names(wong_colors) <- clean_crops
colors_f <- scale_fill_manual(name = "CROPS", values = wong_colors)
colors_c <- scale_color_manual(name = "CROPS", values = wong_colors)
```

### Performance

```{r echo=F}
r2 <- rbind.data.frame(dcorn$R2, dsoy$R2, dwwheat$R2, dalfalfa$R2, dsorghum$R2, dcotton$R2, dhay$R2)
mse <- rbind.data.frame(dcorn$MSE, dsoy$MSE, dwwheat$MSE, dalfalfa$MSE, dsorghum$MSE, dcotton$MSE, dhay$MSE)
dic <- rbind.data.frame(dcorn$DIC, dsoy$DIC, dwwheat$DIC, dalfalfa$DIC, dsorghum$DIC, dcotton$DIC, dhay$DIC)
cpo <- rbind.data.frame(dcorn$CPO, dsoy$CPO, dwwheat$CPO, dalfalfa$CPO, dsorghum$CPO, dcotton$CPO, dhay$CPO)

perf <- cbind.data.frame(clean_crops, r2, mse, dic, cpo)
colnames(perf) <- c("CROP", "R2", "MSE", "DIC", "CPO")

ggplot(data = perf) +
  geom_bar(aes(x = reorder(CROP, desc(R2)), y = R2, fill = CROP), stat = "identity") +
  project_theme +
  labs(title = "R2 across models",
       x = "",
       y = "",
       fill = "") +
  theme(legend.position = "none") +
  colors_f

```

```{r echo=F}
ggplot(data = perf) +
  geom_bar(aes(x = reorder(CROP, desc(MSE)), y = MSE, fill = CROP), stat = "identity") +
  project_theme +
  labs(title = "MSE across models",
       x = "",
       y = "",
       fill = "") +
  theme(legend.position = "none") +
  colors_f
```

Models with smaller DIC preferred to models with larger DIC.  DIC is useful for model comparison, read more p. 170.

```{r echo=F}
ggplot(data = perf) +
  geom_bar(aes(x = reorder(CROP, desc(DIC)), y = DIC, fill = CROP), stat = "identity") +
  project_theme +
  labs(title = "DIC across models",
       x = "",
       y = "",
       fill = "") +
  theme(legend.position = "none") +
  colors_f
```

Conditional predictive ordinate, LOOCV, compares competitive models in terms of predictive performance, with larger values denoting better fit (see 5.6.1, p. 166).

```{r echo=F}
ggplot(data = perf) +
  geom_bar(aes(x = reorder(CROP, desc(CPO)), y = CPO, fill = CROP), stat = "identity") +
  project_theme +
  labs(title = "CPO across models",
       x = "",
       y = "",
       fill = "") +
  theme(legend.position = "none") +
  colors_f
```
 
Performance table:

```{r echo=F}
perf_sub <- perf %>% filter(!CROP %in% c("Cotton", "Sorghum")) %>% arrange(CROP)

perf_sub

```

 
 
### Coefficients per model

```{r echo=F}
plot_coef(mcorn) + ggtitle("Corn")
```

```{r echo=F}
plot_coef(msoy) + ggtitle("Soy")
```

```{r echo=F}
plot_coef(mwwheat) + ggtitle("Winter wheat")
```

```{r echo=F}
plot_coef(malfalfa) + ggtitle("Alfalfa")
```

```{r echo=F}
plot_coef(mhay) + ggtitle("Hay")
```

```{r echo=F}
plot_coef(mcotton) + ggtitle("Cotton")
```

```{r echo=F}
plot_coef(msorghum) + ggtitle("Sorghum")
```

### Coefficients all models


```{r echo=F}
clean_crops <- as.factor(c("Corn", "Soy", "Winter wheat", "Alfalfa", "Hay"))
mlist <- list(mcorn, msoy, mwwheat, malfalfa, mhay)

plot_all_coef <- function(mlist) {
  
  out <- list()
  
  for (i in 1:length(mlist)) {
    
    model <- mlist[[i]]
    coef <- model$summary.fixed[,c(1, 3, 5)] # mean, 0.025, 0.985
    colnames(coef) <- c("MEAN", "LOW", "HIGH")
    coef$VARS <- rownames(coef)
    coef <- coef %>% filter(VARS != "(Intercept)")
    # drop year factors for now and intercept
    # coef <- coef[2:9,]
    coef$CROP <- clean_crops[i]
    out[[i]] <- coef
  
    
    
  }
  
  out <- do.call("rbind.data.frame", out)
  out$VOI <- NA
  out$VOI[out$VARS == "PERC_IRR"] <- "% ag. land irrigated"
  out$VOI[out$VARS == "PERC_AG"] <- "% agricultural land"
  out$VOI[out$VARS == "T_CEC_SOIL"] <- "Cation exchange capacity"
  out$VOI[out$VARS == "T_OC"] <- "Topsoil organic carbon"
  out$VOI[out$VARS == "T_PH_H2O"] <- "Topsoil pH"
  out$VOI[out$VARS == "T_ESP"] <- "Topsoil salinity"
  out$VOI[out$VARS == "factor(YEAR)2009"] <- "y2009"
  out$VOI[out$VARS == "factor(YEAR)2010"] <- "y2010"
  out$VOI[out$VARS == "factor(YEAR)2011"] <- "y2011"
  out$VOI[out$VARS == "factor(YEAR)2012"] <- "y2012"
  out$VOI[out$VARS == "factor(YEAR)2013"] <- "y2013"
  out$VOI[out$VARS == "factor(YEAR)2014"] <- "y2014"
  out$VOI[out$VARS == "factor(YEAR)2015"] <- "y2015"
  out$VOI[out$VARS == "factor(YEAR)2016"] <- "y2016"
  out$VOI[out$VARS == "factor(YEAR)2017"] <- "y2017"
  out$VOI[out$VARS == "factor(YEAR)2018"] <- "y2018"
  
  ggplot(out %>% filter(!VARS %in% c("(Intercept)"))) +
    geom_errorbarh(height = 0, aes(xmin = LOW, xmax = HIGH, 
                                   y = VOI, color = CROP)) +
    project_theme +
    geom_point(aes(x = MEAN, y = VOI, color = CROP), size = 2) +
    geom_vline(xintercept = 0) +
    labs(x = "",
         y = "",
         color = "") +
    colors_c +
  theme(legend.position = "bottom",
        legend.title = element_blank())
  
  
}

plot_all_coef(mlist) + theme(text = element_text(size = 20))

ggsave("./figs/coefficients.png", width = 15, height = 10)
```


Splitting out time:


```{r echo=F}
clean_crops <- as.factor(c("Corn", "Soy", "Winter wheat", "Alfalfa", "Hay"))
mlist <- list(mcorn, msoy, mwwheat, malfalfa, mhay)

plot_all_coef <- function(mlist) {
  
  out <- list()
  
  for (i in 1:length(mlist)) {
    
    model <- mlist[[i]]
    coef <- model$summary.fixed[,c(1, 3, 5)] # mean, 0.025, 0.985
    colnames(coef) <- c("MEAN", "LOW", "HIGH")
    coef$VARS <- rownames(coef)
    coef <- coef %>% filter(VARS != "(Intercept)")
    # drop year factors for now and intercept
    # coef <- coef[2:9,]
    coef$CROP <- clean_crops[i]
    out[[i]] <- coef
  
    
    
  }
  
  out <- do.call("rbind.data.frame", out)
  out$VOI <- NA
  out$VOI[out$VARS == "PERC_IRR"] <- "% ag. land irrigated"
  out$VOI[out$VARS == "PERC_AG"] <- "% agricultural land"
  out$VOI[out$VARS == "T_CEC_SOIL"] <- "Cation exchange capacity"
  out$VOI[out$VARS == "T_OC"] <- "Topsoil organic carbon"
  out$VOI[out$VARS == "T_PH_H2O"] <- "Topsoil pH"
  out$VOI[out$VARS == "T_ESP"] <- "Topsoil sodicity"
  out$VOI[out$VARS == "factor(YEAR)2009"] <- "y2009"
  out$VOI[out$VARS == "factor(YEAR)2010"] <- "y2010"
  out$VOI[out$VARS == "factor(YEAR)2011"] <- "y2011"
  out$VOI[out$VARS == "factor(YEAR)2012"] <- "y2012"
  out$VOI[out$VARS == "factor(YEAR)2013"] <- "y2013"
  out$VOI[out$VARS == "factor(YEAR)2014"] <- "y2014"
  out$VOI[out$VARS == "factor(YEAR)2015"] <- "y2015"
  out$VOI[out$VARS == "factor(YEAR)2016"] <- "y2016"
  out$VOI[out$VARS == "factor(YEAR)2017"] <- "y2017"
  out$VOI[out$VARS == "factor(YEAR)2018"] <- "y2018"
  
  ggplot(out %>% filter(!VARS %in% c("(Intercept)"), 
                        !VOI %in% c("y2009", "y2010", "y2011", "y2012", "y2013", "y2014", "y2015", "y2016", "y2017", "y2018"))) +
    geom_errorbarh(height = 0, aes(xmin = LOW, xmax = HIGH, 
                                   y = reorder(VOI, LOW), color = CROP), position = position_dodge(width = 0.3)) +
    project_theme +
    geom_point(aes(x = MEAN, y = reorder(VOI, LOW), color = CROP), size = 2, alpha = 0.8, 
               position = position_dodge(width = 0.3)) +
    geom_vline(xintercept = 0) +
    labs(x = "",
         y = "",
         color = "") +
    colors_c +
  theme(legend.position = "right",
        legend.title = element_blank()) 
  
  
}

plot_all_coef(mlist) + theme(text = element_text(size = 12))

ggsave("./figs/coefficients_notime.png", width = 7, height = 3)
```


Visualizing time:


```{r echo=F}
clean_crops <- as.factor(c("Corn", "Soy", "Winter wheat", "Alfalfa", "Hay"))
mlist <- list(mcorn, msoy, mwwheat, malfalfa, mhay)

plot_all_coef <- function(mlist) {
  
  out <- list()
  
  for (i in 1:length(mlist)) {
    
    model <- mlist[[i]]
    coef <- model$summary.fixed[,c(1, 3, 5)] # mean, 0.025, 0.985
    colnames(coef) <- c("MEAN", "LOW", "HIGH")
    coef$VARS <- rownames(coef)
    coef <- coef %>% filter(VARS != "(Intercept)")
    # drop year factors for now and intercept
    # coef <- coef[2:9,]
    coef$CROP <- clean_crops[i]
    out[[i]] <- coef
  
    
    
  }
  
  out <- do.call("rbind.data.frame", out)
  out$VOI <- NA
  out$VOI[out$VARS == "PERC_IRR"] <- "% ag. land irrigated"
  out$VOI[out$VARS == "PERC_AG"] <- "% agricultural land"
  out$VOI[out$VARS == "T_CEC_SOIL"] <- "Cation exchange capacity"
  out$VOI[out$VARS == "T_OC"] <- "Topsoil organic carbon"
  out$VOI[out$VARS == "T_PH_H2O"] <- "Topsoil pH"
  out$VOI[out$VARS == "T_ESP"] <- "Topsoil salinity"
  out$VOI[out$VARS == "factor(YEAR)2009"] <- "2009"
  out$VOI[out$VARS == "factor(YEAR)2010"] <- "2010"
  out$VOI[out$VARS == "factor(YEAR)2011"] <- "2011"
  out$VOI[out$VARS == "factor(YEAR)2012"] <- "2012"
  out$VOI[out$VARS == "factor(YEAR)2013"] <- "2013"
  out$VOI[out$VARS == "factor(YEAR)2014"] <- "2014"
  out$VOI[out$VARS == "factor(YEAR)2015"] <- "2015"
  out$VOI[out$VARS == "factor(YEAR)2016"] <- "2016"
  out$VOI[out$VARS == "factor(YEAR)2017"] <- "2017"
  out$VOI[out$VARS == "factor(YEAR)2018"] <- "2018"
  
  outs <- out %>% filter(VOI %in% c("2009", "2010", "2011", "2012", "2013", "2014", "2015", "2016", "2017", "2018"))
  outs$VOI <- as.numeric(as.character(outs$VOI))
  
  ggplot(outs, aes(x = VOI, y = MEAN, group = CROP, color = CROP)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin = LOW, ymax = HIGH), width = 0.2) +
    project_theme +
    labs(x = "",
         y = "Mean estimated effect",
         color = "")+
    colors_c +
  theme(legend.position = "right",
        legend.title = element_blank()) +
    scale_x_continuous(breaks = 2009:2018, labels = 2009:2018)
  
  

  
  
}

plot_all_coef(mlist) + theme(text = element_text(size = 16))

ggsave("./figs/coefficients_time_only.png", width = 10, height = 5)
```



### Nonlinear effects

```{r echo=F}
variable = "GDD"

crops <-crops[!crops %in% c("sorghum", "cotton")]
clean_crops <- clean_crops[!clean_crops %in% c("Sorghum", "Cotton")]

plot_nle <- function(variable) {
  
  out_list <- list()
  for (i in 1:length(crops)) {

    df <- mlist[[i]]$summary.random[[variable]][,c(1,4:6)]
    df$CROPS <- crops[[i]]
    df$CROP_NAME <- clean_crops[[i]]
    out_list[[i]] <- df
    
  }
  nle <- do.call("rbind.data.frame", out_list)
  
  plt <- ggplot(data = nle) + 
    geom_line(aes(x = ID, y = `0.5quant`, group = CROP_NAME)) +
    geom_ribbon(aes(x = ID, y = `0.5quant`, ymin = `0.025quant`, ymax = `0.975quant`, fill = CROP_NAME), alpha = 0.7) +
    project_theme +
    colors_f +
    labs(x = variable, 
         y = "Effect on yield (standardized)") +
    theme(legend.title = element_blank())
  
  return(plt)
  
}

plot_nle("GDD") + xlab("Growing degree days (standardized)")
```


```{r echo=F}
plot_nle("TP") + xlab("Total precipitation (standardized)") + ylab("")
```

Final plots:

```{r echo=F, message=F, warning=F}
library(ggpubr)
gdd <- plot_nle("GDD") + xlab("Growing degree days (standardized)") + theme(text = element_text(size = 20))
tp <- plot_nle("TP") + xlab("Total precipitation (standardized)") + ylab("") + theme(text = element_text(size = 20))

ggarrange(gdd, tp, labels = c("A", "B"), ncol = 2, nrow = 1, common.legend = T, legend = "bottom")

ggsave("./figs/nonlinear_climate.png", plot = last_plot(), width = 16, height = 8)

```

### Aggregate Z-scores

Note that for now, this doesn't deal with missing data for a particular crop.  The second plot shows the number of crops for each county.

```{r echo=F, message=F, warning=F}
out_list <- list()
for (i in 1:length(crops)) {
  train <- prep_data(crops[[i]])[[1]]
  county_sub <- prep_data(crops[[i]])[[2]]
  diff <- compute_diff(mlist[[i]], train, "FRR")
  diff <- cbind.data.frame(diff$GEOID, diff$DIFF, diff$Z)
  colnames(diff) <- c("GEOID", "DIFF", "Z")
  diff$CROP <- crops[[i]]
  out_list[[i]] <- diff
}
out <- do.call("rbind.data.frame", out_list)
out_sum <- out %>% group_by(GEOID) %>% summarize(ZTOTAL = sum(Z),
                                                 N = n())
out_sum <- merge(county, out_sum, by = "GEOID")

ggplot() +
  geom_sf(data = out_sum, color = "transparent", size = 0.05, aes(fill = ZTOTAL)) +
  geom_sf(data = state, color = "grey", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = "transparent") +
  theme_minimal() +
  scale_fill_gradient2(low = "#d8b365", mid = "#f5f5f5", high = "#5ab4ac", midpoint = 0, na.value = "grey50") +
  labs(fill = "",
       title = "Aggregate Z-scores of difference",
       subtitle = "(Built for each region, with county - regional means)")

```

```{r echo=F, message=F, warning=F}
ggplot() +
  geom_sf(data = out_sum, color = "transparent", size = 0.05, aes(fill = N)) +
  geom_sf(data = state, color = "grey", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = "transparent") +
  theme_minimal() +
  scale_fill_gradient2(low = "#d8b365", mid = "#f5f5f5", high = "#5ab4ac", midpoint = 0, na.value = "grey50") +
  labs(fill = "",
       title = "Number of crops cultivated in each county")
```

```{r echo=F, message=F, warning=F}
out_sum$AVG_Z <- out_sum$ZTOTAL/out_sum$N

ggplot() +
  geom_sf(data = out_sum, color = "transparent", size = 0.05, aes(fill = AVG_Z)) +
  geom_sf(data = state, color = "grey", fill = NA, alpha = 0.1) +
  geom_sf(data = frr_shp, size = 1, fill = "transparent") +
  theme_minimal() +
  scale_fill_gradient2(low = "#d8b365", mid = "#f5f5f5", high = "#5ab4ac", midpoint = 0, na.value = "grey50") +
  labs(fill = "",
       title = "Average Z-score",
       subtitle = "Total Z-score divided by number of crops observed in county")
```

